---
title: "Space"
subtitle: "Exercise 12 --- PMAP 8551, Summer 2024"
author: "YOUR NAME HERE"
date: "DATE GOES HERE"
date-format: "long"
format:
  html:
    toc: true
  pdf:
    toc: true
  docx:
    toc: true
---

---

# Task 1: Reflection

Put your reflection here


# Task 2: World map

```{r}
#| label: load-libraries-data
#| warning: false
#| message: false

library(tidyverse)
library(sf)

# Load and clean internet user data
internet_users <- read_csv("data/share-of-individuals-using-the-internet-1990-2015.csv") |>
  # Rename country code column to ISO_A3 so it matches what's in the Natural Earth shapefile
  rename(users = `Individuals using the Internet (% of population) (% of population)`,
         ISO_A3 = Code)

# Load world shapefile from Natural Earth
# https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
world_shapes <- read_sf("data/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp")
```

```{r}
#| label: combine-shapefile-internet-users

# Only look at 2015
users_2015 <- internet_users |>
  filter(Year == 2015)

users_map <- world_shapes |>
  left_join(users_2015, by = "ISO_A3") |>
  filter(ISO_A3 != "ATA") # No internet in Antarctica. Sorry penguins.
```

Make a map showing the proportion of individuals in each country that had access to the internet in 2015.

```{r}
#| label: plot-2015-internet-users

# TODO: Make a map of internet users with ggplot() + geom_sf()
# Important: Remember to consider the map projection
```


# Task 3: Personal map

Draw your own map with your own points. This could be a map of places you've lived, or a map of places you've visited, or a map of places you want to visit. Anything!

The only requirement is that you find an appropriate shapefile (states, counties, world, etc.), collect latitude and longitude data from Google Maps, and plot the points (with or without labels) on a map. Use multiple shapefiles if you want—add roads, rivers, lakes, whatever. Basically follow the code from the example in the section named ["Making your own geocoded data"](https://datavizs24.classes.andrewheiss.com/example/12-example.html#making-your-own-geocoded-data).

```{r}
#| label: personal-map

# TODO: Make a map
```
